Are shade leaves lying in wait? How instantaneous responses of mesophyll conductance to light availability in shade leaves affect theories of optimal canopy carbon gain

### 2015-07-13 manuscript/manuscript_wtc3.Rmd
added results, figures, and new cites

```{r, results='asis', echo=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

@@

where C~c~ is the chloroplast CO~2~ partial pressure. Examples of this approach to measure gas exchange and carbon isotope discrimination are presented in [@evans2013temperature].

@@

where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation]. Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by leaf transpiration. The g~1~ parameter, an index of water cost of carbon assimilation, was generated by fitting observed values of g~s~ into the optimal stomatal conductance model from [@medlyn2012reconciling].

@@

## Sun and Shade leaf resource distribution

Across all measurement campaigns 1m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001).  PPFD was reduced by 78% in the shade (Figure 1). The distribution of H~2~O, via K, was not different across sun and shade leaves (Table 1).  This was a result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types. Leaf nitrogen, on an area basis, was significantly higher in sun leaves than in shade leaves (p<0.001) by ca. 20% (Table1). No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K, E or leaf N either within or across leaf types.

## Leaf photosynthesis rates and photosynthetic capacity

Leaf photosynthesis across time, leaf types, lightson and temp treatment
Leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment, across the wide range of leaf temperatures measured.

classic aci (acc later after gmes)
vcmax and jmax
realtionship of vcmax and photosynthesis with N

## water use
ITE with VPD for all leaf types and temp treatments
leaf 13C with leaf N and what this stands for

## stomatal and mesophyll conductance

@@

@@

**Figure2**. (a) Photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were developed once for all trees, before the initiation of the drought treatment, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf nitrogen~area~ for each chamber, including sun leaves and shade leaves at low light. For (b) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.28 and 0.35, respectively.

```{r wateruse, fig.width=6, fig.height=8, echo=FALSE}
source("master_scripts/wateruse.R")
```

@@

```{r Agmgs, fig.width=6, fig.height=8, echo=FALSE}
source("master_scripts/Agmgs.R")
```

**Figure 4**. The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.  The response of shade leaf physiology to high light was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca. 25min). Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are approximately 95% confidence intervals from the mean.

@@

**Figure 5**. The mean stomatal conductance (a), mesophyll conductance (b) and photosynthesis rate (c) of sun leaves and shade leaves at both low and high light with standard errors.

@@

>>>>

```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

@@

where C~c~ is the chloroplast CO~2~ partial pressure. Examples of this approach to measure gas exchange and carbon isotope discrimination are presented in [@evans2013temperature]. Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular arispaces to the site of carboxylation were then estimated using eq. 7.

@@

where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation]. Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by leaf transpiration. The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear resgression, where K=0.5 [see @duursma2013near].

@@

## Sun and Shade leaf resource distribution
Across all measurement campaigns 1m canopy integrated PPFD was significantly different between sun and shade leaves (P<0.001) and PPFD was reduced by 78% in the shade (Figure 1). The distribution of H~2~O, via K, was not different across sun and shade leaves (Table 1).  This was the result of neither $\Psi_{pd}$, $\Psi_l$ (Table 1) nor E (Table 2) differing between leaf types. Leaf N, on an area basis, was significantly higher in sun leaves than in shade leaves (p<0.001) by ca. 20% (Table1). No effect of the warming treatment was detected with  PPFD, $\Psi_{pd}$, $\Psi_l$, K, E or N~area~ either within or across leaf types.

## Leaf photosynthesis rates and photosynthetic capacity
Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment (P<0.001, Table 2). Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P<0.001). No effect of the warming treatment was detected on rates of A of sun leaves or shade leaves at low or high light. Photosynthesis across all leaf types and treatments did not vary according across the leaf temperatures measured in this study or time of year (Figure S1a).

\
The photsynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (p<0.001 & p=0.02, respectively), as evident from AC~i~ curves from each chamber (Figure 2a). No effect of the warming treatment was detected on either parameter (Table 1). Within individual chambers Vc~max~ was postively related to leaf N~area~ across leaf types and temperature treatments (P =0.01, Figure 2b).  Additionally, leaf N~area~ was postively related to A across gas exchange campaigns and leaf types (P<0.001, Figure 2c).  As Vc~max~ is condisdered implicit with g~m~ (results below) with conventional AC~i~ curves, a AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.  When accounting for the effect of g~m~, photosynthetic capacity was still greater in sun leaves than shade leaves (Figure S2)

## Leaf water use efficiency
Leaf ITE was significantly greater in sun leaves than in shade leaves (P=0.001, Table 2). Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P<0.001).  ITE in sun leaves was reduced in the warming treatment (P=0.021) but no effect was noted in shade leaves with low or high light. For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal ITE model (Figure 3a).  The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.  Across leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure S2a)

Leaf $\delta$^13^C significantly decreased from sun leaves to shade leaves by ca. 2&permil; (p<0.001, Table 1). Within leaf types no affects of the warming treatment on leaf $\delta$^13^C were detected. Leaf $\delta$^13^C and N~area~ were postively correlated for all leaves (P<0.001, Figure 3b).

leaf 13C with leaf N and what this stands for

## Stomatal conductance, mesophyll conductance and CO~2~ drawdown
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%), under their local light environment (P=0.004, Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P<0.001, Figure 4a). In sun leaves, g~s~ and A showed optimal behavior with diminishing rates of A with highest values of g~s~ (Figure 5a).  However, in shade leaves this relationship was clearly uncoupled.  With increased light availability, both increases in A and g~s~ in shade leaves extened the realtionship of A and g~s~ seen in sun leaves.

\
Mean g~m~ was significantly higher in sun compared to shade leaves (+27%), under their local light environment (P=<0.001, Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P<0.001, Figure 4b). Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure 4b,c). Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves in high light resulted in the highest rates of A (Figure 5b). No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types or light treatments. Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures S1b and S2b, respectively) and only a slight negative relationship was detected with g~m~ and increasing leaf temperature.

\
The combination of a non-limiting supply of leaf K and more open stomata lead to a significant increase in C~i~ in shade compared to sun leaves (Figure 6a). As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves. As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordiation between gm and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure 6c).  This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment. CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable acorss the range of temperatures measured and gas exchange campaigns (Figure S2c and S1c, respectively).

## delta contrib, deltav cica

The difference between $\delta$^13^Ce and $\delta$^13^Co was mostly accounted for by discrimination associated with mesophyll conductance (85%) compared to contributions from respiration and photorespiration.

@@

photosynthetic capacity: Mesophyll conductance asymmetrically affects key photosynthetic parameters and an assumption of infinite mesophyll conductance leads to underestimation of the maximum carboxylation rate Vcmax, maximum electron transport rate Jmax [@sun2014asymmetrical].

@@

**Figure 2**. (a) Photosynthetic CO~2~ response (ACi) curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were developed once for all trees, before the initiation of the drought treatment, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf nitrogen~area~ for each chamber, including sun leaves and shade leaves at low light. (c) the relationship between For (b,c) the dashed line represents the significant linear model fit for all leaves with a marginal and conditional r^2^ of 0.28 and 0.35 for (b), and 0.24 and 0.33 for (c).

```{r wateruse, fig.width=6, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE}
source("master_scripts/wateruse.R")
```

@@

@@

**Figure 4**. The mean stomatal conductance (a), mesophyll conductance (b) and photosynthesis rate (c) of sun leaves and shade leaves at both low and high light with standard errors.

```{r Agmgs, fig.width=6, fig.height=8, echo=FALSE}
source("master_scripts/Agmgs.R")
```  
**Figure 5**. The response of leaf photosynthesis rates to stomatal conductance (a) and mesophyll conductance (b) for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments.  The response of shade leaf physiology to high light was recorded once both photosynthesis and stomatal conductance were stable in the gas exchange leaf cuvette (ca. 25min). Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are approximately 95% confidence intervals from the mean.

@@

# Supporting Information {.unnumbered}
```{r gasex_temp1, fig.width=6, fig.height=10, echo=FALSE}
source("master_scripts/gasex_temp.R")
```  
**Figure S1**. Reponse of photosynthesis (a), mesophyll conductance (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.  Shaded symbols represents each monthyl measurement campaign.

```{r gasex_temp2, fig.width=6, fig.height=10, echo=FALSE}
source("master_scripts/gasex_temp2.R")
```  
**Figure S2**. Reponse of vapor pressure deficit (a), stomatal conductance (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.  Shaded symbols represents each monthyl measurement campaign.

```{r acc, fig.width=6, fig.height=6, echo=FALSE}
source("master_scripts/ACc_figure.R")
```

**Figure S3**. Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments. C~c~ values were predicited with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).

```{r deltacica, fig.width=6, fig.height=6, echo=FALSE}
source("master_scripts/delta_cica.R")
```

**Figure S4**.  Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.  The solid line represents the the theortical line for C3 plants from [@evans1986carbon].
